PhD Title \\ PhD Thesis

### 2024-06-14 Latex_files/Thesis/chapters/chapter 2/main.tex
updates chapter 1
updates

@@

Later chapters cover the dynamics of spherical aggregates and anisotropic scatterers, these subjects are particularly difficult to characterise using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}. As an example consider a symmetric dimer as a paradigmatic aggregate; if we consider the Langevin equation for such a aggregate within an optical trap we have:
\begin{align}
\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)
\end{align}
Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational motion is generalised to a 3 dimensional case. Except now, the dimer is undergoing random rotational motion in addition to its Brownian translational motion the first term on the right hand side is no longer purely a function of the dimer's position but also on its orientation. The rotational form of the Langevin equation for a dipole within an external potential is given as:
\begin{align}
  \frac{{d}\vec{u}(t)}{{dt}}
  =
  \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
  + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}
Where $\vec{u}(t)$ is the unit vector aligned along the centres of the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the rotational friction coefficient which is given as $\gamma_R = 8\pi\eta r^3$, if the dimer is within a harmonic potential we can write the first term on the right hand side as $\frac{\vec{\kappa}_u}{\gamma} \times \vec{u}(t)$, where $\vec{\kappa}_u$ is the rotational stiffness vector. $\lambda(t)$ is the Brownian rotations from the surrounding fluid, like in the translational case the Brownian rotations are normally distributed and are also uncorrelated so that:
\begin{align}
  \left<\lambda(t)\lambda(t')\right> = \delta_{ij}\delta(t-t')
\end{align}
For an asymmetric scatterer whose radius is comparable to that of the electric field's wavelength we now have a system of simultaneous equations:
\begin{align}
	\label{eq:full_langevin}
  \frac{{d}\vec{r}(t)}{{dt}}
  &=
    \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2D}\eta(t)
  \\
  \frac{{d}\vec{u}(t)}{{dt}}
  &=
    \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t)
    + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}
Fortunately, we do not need to solve these directly as the latter two
random variables can be easily approximated if the thermal energy of
the system is know, and the rate of change can be assumed as linear if
we take a sufficiently small time step that
$\Delta t~\ll~\kappa_x/\gamma \ \& \ \Delta t \ll
\kappa_u/\gamma_R$. In doing so we now only need to compute the
optical force and torque applied to our dimer, this has already been
compiled in a MATLAB package called \textit{Optical Tweezer Toolbox}
or \textit{ott} \cite{Nieminen2007}. In which they use the results
from \cite{Farsund1996} to compute both the optical force and torque
using the beam coefficients - the form given by
\cite{Crichton2000THEMD} is an easier form to compute:

@@

We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle \textit{ott} uses the t-matrix method, this is not essential for a simple sphere but is essential for complex shaped particles such as dimers. The scattered and incident fields are then combined together in the far field to get $I(x,y)$, the quadrant and overall signals are calculated via:
\begin{align}
	Q_i &= \sum_{n,m} I(x_{i,n}, y_{i,m}) \\
	S_{x} &= \frac{(Q_1+Q_2)-(Q_3+Q_4)}{\sum I_0(x,y)} \\
	S_{y} &= \frac{(Q_1+Q_3)-(Q_2+Q_4)}{\sum I_0(x,y)} \\
	S_{z} &= \frac{(Q_1+Q_2+Q_3+Q_4}{\sum I_0(x,y)}
\end{align}
Where the denominator is the total intensity on the QPD while there is no particle within the trap. The QPD sensitivity is dependent on both the polarisation of the incident beam and the displacement of the target sphere, this is shown in fig.~\ref{fig:totalfield}.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.85\linewidth]{fixed_polarisation.png}
	\captionsetup{margin=0.5cm}
	\caption{Total Field incident on the quadrant photo diode for a several different displacements. Top Left: $2\mu m $ diameter silica sphere placed at the focus of a  beam. Top Right: silica sphere is now displaced by $1 \mu m$ along the x-axis. Bottom Right: silica sphere is now displaced by $1 \mu m$ along the y-axis. Bottom Left: silica sphere is now displaced by $0.707 \mu m$ along the x and y axis. The trapping beam is a $TEM_00$ Gaussian, polarised along the along the x-axis. $S_x$, $S_y$, and $S_z$ are shown to the right of each graph, $Q_1$, $Q_2$, $Q_3$, and $Q_4$ are shown above each graph.}
	\label{fig:totalfield}
\end{figure}

>>>>

\begin{figure}[h]
	\centering
	\includegraphics[width=0.67\linewidth]{finite_differences.png}
	\caption{Example trajectory created using Finite Differences method
			 for a $2\mu m$ diameter sphere. Trap stiffness's were estimated using \textit{ott} at $\kappa_x = \kappa_y = -100\ pN/\mu m$ and $\kappa_z = -25\ pN/\mu m$. The particle's motion can be localised around the shaded ellipsoid.}
\end{figure}

@@

Later chapters cover the dynamics of spherical aggregates and anisotropic scatterers, these subjects are particularly difficult to characterise using conventional calibration techniques \cite{Li2008, Yogesha2011PreciseCO}. As an example consider a symmetric dimer as a paradigmatic aggregate; if we consider the Langevin equation for such a aggregate within an optical trap we have:
\begin{align}
\frac{{d}\vec{r}(t)}{{dt}} = \frac{\vec{\kappa}_x}{\gamma}\vec{r}(t) + \sqrt{2\vec{D}_x}\eta(t)
\end{align}
Where $x(t)$ is replaced with $\vec{r}(t)$ to signify that the translational motion is generalised to a 3 dimensional case. Except now, the dimer is undergoing random rotational motion in addition to its Brownian translational motion the first term on the right hand side is no longer purely a function of the dimer's position but also on its orientation. The rotational form of the Langevin equation for a dipole within an external potential is given as:
\begin{align}
  \frac{{d}\vec{u}(t)}{{dt}}
  =
  \frac{\mu}{\gamma_R}\left[\vec{u}(t)\times E(t)\right]\times \vec{u}(t)
  + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}
Where $\vec{u}(t)$ is the unit vector aligned along the centres of the two spheres, $\mu$ is its dipole moment, and $\gamma_R$ is the rotational friction coefficient which is given as $\gamma_R = 8\pi\eta r^3$ for a sphere, if the dimer is within a harmonic potential we can write the first term on the right hand side as $\frac{\vec{\kappa}_u}{\gamma} \times \vec{u}(t)$, where $\vec{\kappa}_u$ is the rotational stiffness vector. $\lambda(t)$ is the Brownian rotations from the surrounding fluid, like in the translational case the Brownian rotations are normally distributed and are also uncorrelated so that:
\begin{align}
  \left<\lambda(t)\lambda(t')\right> = \delta_{ij}\delta(t-t')
\end{align}
For an asymmetric scatterer whose radius is comparable to that of the electric field's wavelength we now have a system of simultaneous equations:
\begin{align}
	\label{eq:full_langevin}
  \frac{{d}\vec{r}(t)}{{dt}}
  &=
    \frac{\vec{\kappa}_x(\vec{u}(t))}{\gamma}\vec{r}(t) + \sqrt{2D}\eta(t)
  \\
  \frac{{d}\vec{u}(t)}{{dt}}
  &=
    \frac{\vec{\kappa}_u(\vec{r}(t))}{\gamma_R}\times \vec{u}(t)
    + \sqrt{2\vec{D}_R}\lambda(t)\times \vec{u}(t)
\end{align}
Fortunately, we do not need to solve these directly as the latter two
random variables can be easily approximated if the thermal energy of
the system is know, and the rate of change can be assumed as linear if
we take a sufficiently small time step that
$\Delta t~\ll~\kappa_x/\gamma \ \& \ \Delta t \ll
\kappa_u/\gamma_R$. In doing so we now only need to compute the
optical force and torque applied to our dimer, this has already been
compiled in a MATLAB package called \textit{Optical Tweezer Toolbox}
or \textit{ott} \cite{Nieminen2007}. In which they use the results
from \cite{Farsund1996} to compute both the optical force and torque
using the beam coefficients - the form given by
\cite{Crichton2000THEMD} is an easier form to compute:

@@

We can if we want further restrict the incident beam by applying setting the truncation angle to match our microscope object, this essentially applies a cut off point to the In order to compute the scattering from the target particle \textit{ott} uses the t-matrix method, this is not essential for a simple sphere but is essential for complex shaped particles such as dimers. The scattered and incident fields are then combined together in the far field to get $I(x,y)$, the quadrant and overall signals are calculated via:
\begin{align}
	Q_i &= \sum_{n,m} I(x_{i,n}, y_{i,m}) \\
	S_{x} &= \frac{(Q_1+Q_2)-(Q_3+Q_4)}{\sum I_0(x,y)} \\
	S_{y} &= \frac{(Q_1+Q_3)-(Q_2+Q_4)}{\sum I_0(x,y)} \\
	S_{z} &= \frac{(Q_1+Q_2+Q_3+Q_4}{\sum I_0(x,y)}
\end{align}
Where the denominator is the total intensity on the QPD while there is no particle within the trap. The QPD sensitivity is dependent on both the polarisation of the incident beam and the displacement of the target sphere, this is shown in fig.~\ref{fig:totalfield}.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.85\linewidth]{fixed_polarisation.png}
	\captionsetup{margin=0.5cm}
	\caption{Total Field incident on the quadrant photo diode for a several different displacements. Top Left: $2\mu m $ diameter silica sphere placed at the focus of a  beam. Top Right: silica sphere is now displaced by $1 \mu m$ along the x-axis. Bottom Right: silica sphere is now displaced by $1 \mu m$ along the y-axis. Bottom Left: silica sphere is now displaced by $0.707 \mu m$ along the x and y axis. The trapping beam is a $TEM_{00}$ Gaussian, polarised along the along the x-axis. $S_x$, $S_y$, and $S_z$ are shown to the right of each graph, $Q_1$, $Q_2$, $Q_3$, and $Q_4$ are shown above each graph.}
	\label{fig:totalfield}
\end{figure}

### 2024-06-14 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

Where the ellipticity of the beam is given by $e = (1-\kappa_y/\kappa_x)^{0.5}$ and is a measure of the symmetry of 
the beam wavefront. Its clear from these initial results that the 
QPD is more sensitive to changes along the y-axis than the x-axis 
when compared to the direct \textit{ott} calculations. Typically,
even an industrial Gaussian beam will produce an elliptical diffraction
limit spot when heavily focused; in their tutorial for optimizing 
the PSD analysis, Berg and Sorensen reported a ellipticity of
around 15 \% after a total calibration time of 81 seconds \cite{BergSoerensen2004}. If we now consider a symmetric the 
discrepancy between \textit{ott} and the simulated QPD grows.

@@

Now we see that the \textit{ott} predicts a more elliptical 
trap compared to the QPD model which says the trap is far
more symmetrical while trapping a symmetric dimer. This 
change can be partially explained by the fact that rotational
effects are not accounted for in the Lorentzian power spectra,
only translational motion. Typically, rotational motion is only
ever detected when it is periodic (take for example a power spectra
from a rotating birefringent sphere), when the motion is stochastic
the entire power spectra is effected making it near impossible
to separate the translational and rotational contributions from a 
single power spectra.

>>>>

Where the ellipticity of the beam is given by $e = (1-\kappa_y/\kappa_x)^{0.5}$ and is a measure of the symmetry of 
the beam wavefront. Its clear from these initial results that the 
QPD is more sensitive to changes along the y-axis than the x-axis 
when compared to the direct \textit{ott} calculations. Typically,
even an industrial Gaussian beam will produce an elliptical diffraction
limit spot when heavily focused; in their tutorial for optimizing 
the PSD analysis, Berg and Sorensen reported a ellipticity of
around 15 \% after a total calibration time of 81 seconds \cite{BergSoerensen2004}. The reason for this discrepancy
can be explained partly by the fact that our estimation of the 
trap geometry via \textit{ott} is based on the the actual 
interactions with the beam near close to the focus whereas
a QPD is estimating the trap geometry by extrapolating from
far field scattering signals. If we now consider a symmetric the 
discrepancy between \textit{ott} and the simulated QPD grows.

@@

Now we see that the \textit{ott} predicts a more elliptical 
trap compared to the QPD model which says the trap is far
more symmetrical while trapping a symmetric dimer. The reason that
\textit{ott} no longer expects a circular trap could be due to
how it computes the beam shape coefficients; by point matching in 
the far field before the focus means a loss in accuracy for objects
that trap above the focus. The change in the QPD estimation  
can be partially explained by the fact that rotational
effects are not accounted for in the Lorentzian power spectra,
only translational motion. Typically, rotational motion is only
ever detected when it is periodic (take for example a power spectra
from a rotating birefringent sphere), when the motion is stochastic
the entire power spectra is effected making it near impossible
to separate the translational and rotational contributions from a 
single power spectra.
